#------------------------- Statistics for CAPs in CPP Publication -------------------------#

Salome Häuselmann (salome.haeuselmann@gmail.com), Winter 2025

This script visualises the Temporal Characteristics (TCs) of the CAPs and calculates statistics. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setting up R

## Clear Workspace and Load Packages

```{r}
rm(list = ls())

pacman::p_load(tidyverse, dplyr, xlsx, readxl, ggplot2, ggpubr, forcats, cowplot, 
               plyr, rstatix, reshape2, writexl, car, lawstat, multcomp, tidyr,
               gtsummary, gridExtra, gt,lmerTest, MuMIn, gg.gap, lme4, Gmisc, 
               lubridate, readxl, openxlsx, dplyr, multcomp, effectsize, car, dplyr, purrr)
```


## Set Working Directory

```{r}
# Seed directory
file_path <- file.path("")

```

## Read in Excel file

```{r}
Duration <- read.xlsx(file_path, sheet = "Duration")
ReOccurence <- read.xlsx(file_path, sheet = "RelativeOccurence")
Entries <- read.xlsx(file_path, sheet = "Entries")
ReEntries_t <- read.xlsx(file_path, sheet = "RelativeEntries_t")
ReEntries_r <- read.xlsx(file_path, sheet = "RelativeEntries_r")
Transition_CPP <- read.xlsx(file_path, sheet = "Transition_CPP", colNames = FALSE)
Transition_HC <- read.xlsx(file_path, sheet = "Transition_HC", colNames = FALSE)
ExcludedFrames <- read.xlsx(file_path, sheet = "ExcludedFrames")
SelectedFrames <- read.xlsx(file_path, sheet = "SelectedFrames")
RetainedFrames <- read.xlsx(file_path, sheet = "RetainedFrames")
RelRetainedFrames <- read.xlsx(file_path, sheet = "RelRetainedFrames")
CRF<- read.xlsx("")

```

# Reorganize Data

## Add Clinical Characteristics to the Temporal Characteristics

```{r}
# Correct the P-Code Format
CRF <- dplyr::rename(CRF, ID = p_code)

# Rename "value" of the TC and combine all TC in one sheet.
head(Duration) 
Duration <- dplyr::rename(Duration, duration_value = value)
ReOccurence <- dplyr::rename(ReOccurence, ReOccurence_value = value)
Entries <- dplyr::rename(Entries, entries_value = value)
ReEntries_t <- dplyr::rename(ReEntries_t, reentries_t_value = value)
ReEntries_r <- dplyr::rename(ReEntries_r, reentries_r_value = value)
ExcludedFrames <- dplyr::rename(ExcludedFrames, excludedframes_value = value)
SelectedFrames <- dplyr::rename(SelectedFrames, selectedframes_value = value)
RetainedFrames <- dplyr::rename(RetainedFrames, retainedframes_value = value)
RetainedFrames <- RetainedFrames %>%
  left_join(CRF %>% select(ID, group), by = "ID")
RelRetainedFrames <-dplyr::rename(RelRetainedFrames, relretainedframes_value = value)


df_TC <- merge(Duration, ReOccurence, by = c("ID", "State"), all = FALSE)
df_TC <- merge(df_TC, Entries, by = c("ID", "State"), all = FALSE)
df_TC <- merge(df_TC, ReEntries_t, by = c("ID", "State"), all = FALSE)
df_TC <- merge(df_TC, ReEntries_r, by = c("ID", "State"), all = FALSE)

df_TC_CRF <- df_TC %>%
  left_join(CRF, by = "ID") %>%
  left_join(ExcludedFrames %>% select(ID, excludedframes_value), by = "ID") %>%
  left_join(SelectedFrames %>% select(ID, selectedframes_value), by = "ID") %>%
  left_join(RetainedFrames %>% select(ID, retainedframes_value), by = "ID") %>%
  left_join(RelRetainedFrames %>% select(ID, relretainedframes_value), by = "ID")
  
# save df subset
new_subfolder <- file.path(output_directory, "for_cor")

# Create the new subfolder if it doesn’t already exist
if (!dir.exists(new_subfolder)) {
  dir.create(new_subfolder)
}

output_file <- file.path(new_subfolder, "AnteriorInsula_K3_for_corr.xlsx")
write.xlsx(df_TC_CRF, output_file) # This can now be used for correlation analysis. 

df_TC_CRF$State<-revalue(df_TC_CRF$State, c("CAP 1"="CAP1","CAP 2"="CAP2","CAP 3"="CAP3"))

# Show the class of selected variables
vars <- c("ID","group","State","gender","psychMed","pain_med",
          "age","sum_stai2_bdi","selectedframes_value",
          "duration_value","reentries_r_value","ReOccurence_value")

sapply(df_TC_CRF[vars], class)

# Adapt classes
df_TC_CRF$ID     <- factor(df_TC_CRF$ID)
df_TC_CRF$group  <- factor(df_TC_CRF$group)
df_TC_CRF$State  <- factor(df_TC_CRF$State)
df_TC_CRF$gender <- factor(df_TC_CRF$gender)

# If psychMed is a 0/1 indicator or categorical, treat as factor:
df_TC_CRF$psychMed <- factor(df_TC_CRF$psychMed)
df_TC_CRF$pain_med <- factor(df_TC_CRF$pain_med)

```

## Check the data

```{r}
str(df_TC_CRF)
names(df_TC_CRF)

```



## Violin-Plots + Statistics 

### TC Visualization + multiple t-test FDR corrected

#### Duration

```{r}
d <- ggplot(df_TC_CRF, aes(x = group, y = duration_value, fill = group)) +   
  geom_violin(trim = FALSE) +   
  facet_wrap(~ State) +  # Create separate plots for each State
  geom_jitter(aes(color = group), width = 0.2, size = 0.8, alpha = 0.6, show.legend = FALSE) +
  theme_minimal() +   
 scale_fill_manual(values = c("#7E7EBF", "#F5B1B4"))+ 
  scale_color_manual(values = c("#160A78", "#EF6472")) +
    scale_y_continuous(limits = c(0, 5), expand = c(0, 0), breaks = seq(0, 5, 1)) +
  labs(y = "average duration [s]") +   
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),  
    axis.text.y = element_text(size = 15),                         
    axis.title.x = element_text(size = 15),                        
    axis.title.y = element_text(size = 15),                        
    plot.title = element_text(size = 15),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(), 
    strip.text = element_text(size = 15),  
    axis.line = element_line(color = "black"),    
    axis.ticks.y = element_line(color = "black", linewidth = 0.3),
    axis.ticks.length = unit(3, "pt")    
  )

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}

fig3a <- d + stat_summary(fun.data=data_summary)

# Save plot
#plot.new()
#tiff("Duration_Amy.tiff", height = 360, width = 500, units='mm' ,
#     compression = "lzw", res = 300)    

# dev.off() # Turn the PDF device off
ggsave("Figure_3a.tiff",
       plot = fig3a,
       dpi = 600,
       width = 5,
       height = 6,
       units = "in",
       device = "tiff",
       compression = "lzw")

# Summary Statistics
df_TC_CRF %>%
  group_by(group,State) %>%
  get_summary_stats(duration_value, type = "mean_sd")

```

#### Relative Entries (normalized to retained frames)

```{r}
e <- ggplot(df_TC_CRF, aes(x = group, y = reentries_r_value, fill = group)) +
  geom_violin(trim = FALSE) +
  facet_wrap(~ State) +  # Create separate plots for each State
  geom_jitter(aes(color = group), width = 0.2, size = 0.8, alpha = 0.6, show.legend = FALSE) +
  theme_minimal() +
  scale_fill_manual(values = c("#7E7EBF", "#F5B1B4"))+ 
  scale_color_manual(values = c("#160A78", "#EF6472")) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0), breaks = seq(0, 1, 0.2)) + # breaks at 0, 0.2, 0.4, 0.6, 0.8, 1
  labs(y = "relative number of entries") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13),  
    axis.text.y = element_text(size = 15),                         
    axis.title.x = element_text(size = 15),                        
    axis.title.y = element_text(size = 15),                       
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(), 
    strip.text = element_text(size = 15),  
    axis.line = element_line(color = "black"),
    axis.ticks.y = element_line(color = "black", linewidth = 0.3),
    axis.ticks.length = unit(3, "pt")   
  )

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}

fig3b <- e + stat_summary(fun.data=data_summary)
fig3b
ggsave("Figure_3b.tiff",
       plot = fig3b,
       dpi = 600,
       width = 5,
       height = 6,
       units = "in",
       device = "tiff",
       compression = "lzw")

# Summary Statistics
df_TC_CRF %>%
  group_by(group,State) %>%
  get_summary_stats(reentries_r_value, type = "mean_sd")

```

#### Relative Occurence

```{r}
r <- ggplot(df_TC_CRF, aes(x = group, y = ReOccurence_value, fill = group)) +
  geom_violin(trim = FALSE) +
  facet_wrap(~ State) +  # Create separate plots for each State
  geom_jitter(aes(color = group), width = 0.2, size = 0.8, alpha = 0.6, show.legend = FALSE) +
  theme_minimal() +
  scale_fill_manual(values = c("#7E7EBF", "#F5B1B4"))+ 
  scale_color_manual(values = c("#160A78", "#EF6472")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +  
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0), breaks = seq(0, 1, 0.2)) + 
  labs(y = "relative occurrence") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 13), 
    axis.text.y = element_text(size = 15),                         
    axis.title.x = element_text(size = 15),                        
    axis.title.y = element_text(size = 15),                       
    plot.title = element_text(size = 15),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(), 
    strip.text = element_text(size = 15),  
        axis.line = element_line(color = "black"), 
    axis.ticks.y = element_line(color = "black", linewidth = 0.3),
    axis.ticks.length = unit(3, "pt")    
  )

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}

fig3c <- r + stat_summary(fun.data=data_summary)

ggsave("Figure_3c.tiff",
       plot = fig3c,
       dpi = 600,
       width = 5,
       height = 6,
       units = "in",
       device = "tiff",
       compression = "lzw")
# Summary Statistics
df_TC_CRF %>%
  group_by(group,State) %>%
  get_summary_stats(ReOccurence_value, type = "mean_sd")

```

### TC Modelling Primary Models
#### Duration

```{r}
### Primary Model 
## post-hoc with the linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_duration_Mx<-with(df_TC_CRF, lmer(duration_value ~ groups + gender + age + sum_stai2_bdi + psychMed +  selectedframes_value + (1 | ID)))
summary(model_posthoc_duration_Mx)

posthoc_results_duration_Mx <- summary(glht(model_posthoc_duration_Mx, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))
posthoc_results_duration_Mx

```

#### Relative Entries (normalized to retained frames)

```{r}
## post-hoc for linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_entries_Mx<-with(df_TC_CRF, lmer(reentries_r_value ~ groups + gender + age + sum_stai2_bdi + psychMed + selectedframes_value+ (1 | ID)))
summary(model_posthoc_entries_Mx)


posthoc_results_entries_Mx <- summary(glht(model_posthoc_entries_Mx, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))

posthoc_results_entries_Mx

```

#### Relative Occurence

```{r}
## post-hoc for linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_RO_Mx<-with(df_TC_CRF, lmer(ReOccurence_value ~ groups + gender + age + sum_stai2_bdi + psychMed  + selectedframes_value+ (1 | ID)))
summary(model_posthoc_RO_Mx)

posthoc_results_RO_Mx <- summary(glht(model_posthoc_RO_Mx, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))

posthoc_results_RO_Mx
```


#### FDR over the 3 models

```{r}
# make sure your factors are prepared as you did

# your contrasts (add CAP4 lines back in if you want them included in the global FDR)
K <- c("(HC.CAP1)-(CPP.CAP1)=0",
       "(HC.CAP2)-(CPP.CAP2)=0",
       "(HC.CAP3)-(CPP.CAP3)=0")

# helper to extract raw p-values from a model’s glht
glht_table <- function(model, K, model_name) {
  smry <- summary(glht(model, linfct = mcp(groups = K)), test = adjusted("none"))
  print(smry)
  data.frame(
    model     = model_name,
    contrast  = names(smry$test$coefficients),
    estimate  = as.numeric(smry$test$coefficients),
    se        = as.numeric(smry$test$sigma),
    z_value   = as.numeric(smry$test$tstat),
    p_raw     = as.numeric(smry$test$pvalues),
    stringsAsFactors = FALSE
  )
}

# gather all posthoc tests from all models
all_tests <- bind_rows(
  glht_table(model_posthoc_entries_Mx,  K, "entries"),
  glht_table(model_posthoc_duration_Mx, K, "duration"),
  glht_table(model_posthoc_RO_Mx,       K, "reoccurrence")
)

# add FDR corrections
all_tests <- all_tests %>%
  mutate(
    p_fdr_global = p.adjust(p_raw, method = "fdr")               # FDR across ALL models/contrasts
  ) %>%
  group_by(model) %>%
  mutate(
    p_fdr_within_model = p.adjust(p_raw, method = "fdr")         # optional: FDR within each model
  ) %>%
  ungroup() %>%
  arrange(p_fdr_global)

# inspect results
all_tests
# If you want a per-model table:
split(all_tests, all_tests$model)

####Effect size calculation: Cohens d' = difference in mean/ SD_pooled (SD of residuals)

all_tests <- all_tests %>%
  mutate(
    cohens_d = case_when(
      model == "entries" ~ estimate / sigma(model_posthoc_entries_Mx),
      model == "duration" ~ estimate / sigma(model_posthoc_duration_Mx),
      model == "reoccurrence" ~ estimate / sigma(model_posthoc_RO_Mx)
    )
  )

# View results with effect sizes
all_tests %>%
  select(model, contrast, estimate, se, p_raw, p_fdr_global, cohens_d)

```

### TC Modelling Minimal Models

#### Duration

```{r}
### Minimal Model 
## post-hoc with the linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_duration_Mx_min<-with(df_TC_CRF, lmer(duration_value ~ groups + gender + age +  selectedframes_value + (1 | ID)))
summary(model_posthoc_duration_Mx_min)

posthoc_results_duration_Mx_min <- summary(glht(model_posthoc_duration_Mx_min, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))
posthoc_results_duration_Mx_min

```

#### Relative Entries (normalized to retained frames)

```{r}
## post-hoc for linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_entries_Mx_min<-with(df_TC_CRF, lmer(reentries_r_value ~ groups + gender + age + selectedframes_value+ (1 | ID)))
summary(model_posthoc_entries_Mx_min)


posthoc_results_entries_Mx_min <- summary(glht(model_posthoc_entries_Mx_min, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))

posthoc_results_entries_Mx_min

```

#### Relative Occurence

```{r}
## post-hoc for linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_RO_Mx_min<-with(df_TC_CRF, lmer(ReOccurence_value ~ groups + gender + age + selectedframes_value+ (1 | ID)))
summary(model_posthoc_RO_Mx_min)

posthoc_results_RO_Mx_min <- summary(glht(model_posthoc_RO_Mx_min, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))

posthoc_results_RO_Mx_min
```


#### FDR over the 3 models

```{r}
# make sure your factors are prepared as you did

# your contrasts (add CAP4 lines back in if you want them included in the global FDR)
K <- c("(HC.CAP1)-(CPP.CAP1)=0",
       "(HC.CAP2)-(CPP.CAP2)=0",
       "(HC.CAP3)-(CPP.CAP3)=0")

# helper to extract raw p-values from a model’s glht
glht_table <- function(model, K, model_name) {
  smry <- summary(glht(model, linfct = mcp(groups = K)), test = adjusted("none"))
  print(smry)
  data.frame(
    model     = model_name,
    contrast  = names(smry$test$coefficients),
    estimate  = as.numeric(smry$test$coefficients),
    se        = as.numeric(smry$test$sigma),
    z_value   = as.numeric(smry$test$tstat),
    p_raw     = as.numeric(smry$test$pvalues),
    stringsAsFactors = FALSE
  )
}

# gather all posthoc tests from all models
all_tests <- bind_rows(
  glht_table(model_posthoc_entries_Mx_min,  K, "entries"),
  glht_table(model_posthoc_duration_Mx_min, K, "duration"),
  glht_table(model_posthoc_RO_Mx_min,       K, "reoccurrence")
)

# add FDR corrections
all_tests <- all_tests %>%
  mutate(
    p_fdr_global = p.adjust(p_raw, method = "fdr")               # FDR across ALL models/contrasts
  ) %>%
  group_by(model) %>%
  mutate(
    p_fdr_within_model = p.adjust(p_raw, method = "fdr")         # optional: FDR within each model
  ) %>%
  ungroup() %>%
  arrange(p_fdr_global)

# inspect results
all_tests
# If you want a per-model table:
split(all_tests, all_tests$model)

####Effect size calculation: Cohens d' = difference in mean/ SD_pooled (SD of residuals)

all_tests <- all_tests %>%
  mutate(
    cohens_d = case_when(
      model == "entries" ~ estimate / sigma(model_posthoc_entries_Mx_min),
      model == "duration" ~ estimate / sigma(model_posthoc_duration_Mx_min),
      model == "reoccurrence" ~ estimate / sigma(model_posthoc_RO_Mx_min)
    )
  )

# View results with effect sizes
all_tests %>%
  select(model, contrast, estimate, se, p_raw, p_fdr_global, cohens_d)

```

### TC Modelling Extended Models

#### Duration

```{r}
### extended Model 
## post-hoc with the linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_duration_Mx_ex<-with(df_TC_CRF, lmer(duration_value ~ groups + gender + age + sum_stai2_bdi + psychMed + pain_med + selectedframes_value + (1 | ID)))
summary(model_posthoc_duration_Mx_ex)

posthoc_results_duration_Mx_ex <- summary(glht(model_posthoc_duration_Mx_ex, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))
posthoc_results_duration_Mx_ex

```

#### Relative Entries (normalized to retained frames)

```{r}
## post-hoc for linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_entries_Mx_ex<-with(df_TC_CRF, lmer(reentries_r_value ~ groups + gender + age + sum_stai2_bdi + psychMed + pain_med + selectedframes_value+ (1 | ID)))
summary(model_posthoc_entries_Mx_ex)


posthoc_results_entries_Mx_ex <- summary(glht(model_posthoc_entries_Mx_ex, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))

posthoc_results_entries_Mx_ex

```

#### Relative Occurence

```{r}
## post-hoc for linear mixed model ##
# Create interaction term 
df_TC_CRF$groups <- interaction(df_TC_CRF$group, df_TC_CRF$State)
model_posthoc_RO_Mx_ex<-with(df_TC_CRF, lmer(ReOccurence_value ~ groups + gender + age + sum_stai2_bdi + psychMed  + pain_med + selectedframes_value+ (1 | ID)))
summary(model_posthoc_RO_Mx_ex)

posthoc_results_RO_Mx_ex <- summary(glht(model_posthoc_RO_Mx_ex, 
                                linfct = mcp(groups = 
                                               c("(HC.CAP1)-(CPP.CAP1)=0", 
                                                 "(HC.CAP2)-(CPP.CAP2)=0", 
                                                 "(HC.CAP3)-(CPP.CAP3)=0"))), 
                           test = adjusted("fdr"))

posthoc_results_RO_Mx_ex
```


#### FDR over the 3 models

```{r}
# make sure your factors are prepared as you did

# your contrasts (add CAP4 lines back in if you want them included in the global FDR)
K <- c("(HC.CAP1)-(CPP.CAP1)=0",
       "(HC.CAP2)-(CPP.CAP2)=0",
       "(HC.CAP3)-(CPP.CAP3)=0")

# helper to extract raw p-values from a model’s glht
glht_table <- function(model, K, model_name) {
  smry <- summary(glht(model, linfct = mcp(groups = K)), test = adjusted("none"))
  print(smry)
  data.frame(
    model     = model_name,
    contrast  = names(smry$test$coefficients),
    estimate  = as.numeric(smry$test$coefficients),
    se        = as.numeric(smry$test$sigma),
    z_value   = as.numeric(smry$test$tstat),
    p_raw     = as.numeric(smry$test$pvalues),
    stringsAsFactors = FALSE
  )
}

# gather all posthoc tests from all models
all_tests <- bind_rows(
  glht_table(model_posthoc_entries_Mx_ex,  K, "entries"),
  glht_table(model_posthoc_duration_Mx_ex, K, "duration"),
  glht_table(model_posthoc_RO_Mx_ex,       K, "reoccurrence")
)

# add FDR corrections
all_tests <- all_tests %>%
  mutate(
    p_fdr_global = p.adjust(p_raw, method = "fdr")               # FDR across ALL models/contrasts
  ) %>%
  group_by(model) %>%
  mutate(
    p_fdr_within_model = p.adjust(p_raw, method = "fdr")         # optional: FDR within each model
  ) %>%
  ungroup() %>%
  arrange(p_fdr_global)

# inspect results
all_tests
# If you want a per-model table:
split(all_tests, all_tests$model)

####Effect size calculation: Cohens d' = difference in mean/ SD_pooled (SD of residuals)

all_tests <- all_tests %>%
  mutate(
    cohens_d = case_when(
      model == "entries" ~ estimate / sigma(model_posthoc_entries_Mx_ex),
      model == "duration" ~ estimate / sigma(model_posthoc_duration_Mx_ex),
      model == "reoccurrence" ~ estimate / sigma(model_posthoc_RO_Mx_ex)
    )
  )

# View results with effect sizes
all_tests %>%
  select(model, contrast, estimate, se, p_raw, p_fdr_global, cohens_d)

```


#### Retained Frames

```{r}
d <- ggplot(RetainedFrames, aes(x = group, y = retainedframes_value/300, fill = group)) +   
  geom_violin(trim = FALSE) +   
  geom_jitter(aes(color = group), width = 0.2, size = 1.5, alpha = 0.6, show.legend = FALSE) +
  theme_minimal() +   
  scale_fill_manual(values = c("#C0B5BE","#C0B5BE")) + 
  scale_color_manual(values = c("#a08a9a","#a08a9a")) +
  labs(title = "Anterior Insula Activation", x = "Group", y = "temporal fraction") +   
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 15),  
    axis.text.y = element_text(size = 15),                         
    axis.title.x = element_text(size = 15),                        
    axis.title.y = element_text(size = 15),                        
    plot.title = element_text(size = 15),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(), 
        axis.line = element_line(color = "black"), 
    axis.ticks.y = element_line(color = "black", linewidth = 0.3),
    axis.ticks.length = unit(3, "pt")    
  )

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}

d + stat_summary(fun.data=data_summary)

fig2c <- d + stat_summary(fun.data=data_summary)

ggsave("Figure_2c.tiff",
       plot = fig2c,
       dpi = 600,
       width = 5.3,
       height = 6,
       units = "in",
       device = "tiff",
       compression = "lzw")

# Summary Statistics
RetainedFrames %>%
  group_by(group) %>%
  get_summary_stats(retainedframes_value, type = "mean_sd")

# Inspect distribution
ggplot(RetainedFrames, aes(x = retainedframes_value, fill = group)) +
  geom_density(alpha = 0.5) +
  labs(title = "Age distribution by group")

# Test normality (Shapiro-Wilk test) in each group
shapiro.test(RetainedFrames$retainedframes_value[RetainedFrames$group == "CPP"]) 
shapiro.test(RetainedFrames$retainedframes_value[RetainedFrames$group == "HC"]) 

# Check for variance equality (optional)
var.test(retainedframes_value ~ group, data = RetainedFrames)

# Choose the appropriate test
# If both groups are approximately normal and variances are equal -> t-test
# If variances differ -> Welch's t-test
# If non-normal -> Wilcoxon test

# t-tests 
stat.test <- RetainedFrames %>%
  wilcox_test(retainedframes_value ~ group) %>%
  add_significance()
stat.test

wilcox.test(retainedframes_value ~ group, data = RetainedFrames, exact = FALSE)

```